---
title: "Data-preprocessing-and-cell-annotation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data-preprocessing-and-cell-annotation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# 1 Data preprocessing

We used snRNA-seq in mouse jaw masticatory muscles-tendon-bone E12-E16 scMutiome for the preprocessing demo, and the preprocessing demo for snATAC was placed in [Dynamic-network](./vignettes/Dynamic-network.Rmd)

## 1.1 set R library loaction and load packages

```{r echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
.libPaths('~/R/x86_64-pc-linux-gnu-library/4.3/')
library(ggplot2)
library(devtools)
library(tidyverse)
library(pheatmap)
```

## 1.2 load LineageNet

```{r echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
library(LineageNet)
```

## 1.3 set color

```{r eval = TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
my_pal_science6 = c("#A9CA70","#C5D6F0","#F18C54",
                    "#9B59B6","#F7CAC9","#76D7C4")
```

## 1.4 load test data

We gave packing Seurat data for a mouse jaw masticatory muscles-tendon-bone E12-E16 scMutiome

```{r eval = TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
data("bad_genes")

seob = readRDS("/mnt/data/home/tycloud/workspace/scATAC/data/E12_E16_WT_Mand_RPCA_scRNA.rds")

```

## 1.5 Data quality control, dimensionality reduction, clustering

```{r eval = TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
seob = RunSeurat(seob = seob,
                 split_by="sample",
                 integrated="RPCA",
                 outfeatures = bad_genes,
                 coln_add = "all",
                 cellid=NULL,
                 nfeatures=3000,
                 min_cells = 3,
                 min_features = 200,
                 npcs=20,
                 resolution=0.15,
                 k_anchor=3,
                 k_weight = 200,
                 use_old_RNA = F,
                 use_old_Seurat = F)
```

# 2 cell-annotation

## 2.1 plot marker

```{r eval = FALSE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
plotDot(object = seob,
        features=unique(c('Sox9', 'Col2a1',
                          'Pecam1','Cdh5',
                          'Krt14','Epcam',
                          'Plp1','Sox10',
                          'Mitf',
                          'Prrx1', 'Twist1',
                          'Myod1','Myog')),
        mode=1, ID=c('celltype'),
        out_ID = c('celltype'),
        my_pal = colorRampPalette(c(my_pal_science6))(7),
        my_mal_gradientn = c("#4C3399","#F5F2EF","#FFE714"),
        cluster_ID = F,
        cluster_feature = F,
        x_angle=45)
```

## 2.2 plot time and celltype

```{r eval = TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
DimPlot(seob, reduction = "umap", group.by = "sample",label = T)+
  scale_color_manual(values = colorRampPalette(c("#F7903D", "#4D85BD", "#59A95A"))(5))
DimPlot(seob, reduction = "umap", group.by = "celltype",label = T)+
  scale_color_manual(values = colorRampPalette(c(my_pal_science6))(7))
```

# 3 monocle3 (optimal)

```{r eval = FALSE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
cds = Runmonocl3(seob = seob,
                 cluster="seurat_clusters",
                 ncenter = 500)
cds <- order_cells(cds)
```

# 3.1 plot

```{r eval = FALSE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
plot_cells(cds,
           color_cells_by = "seurat_clusters",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)+
  scale_color_manual(values = colorRampPalette(c(my_pal_science6))(length(unique(seob@meta.data$seurat_clusters))))

plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

```
