---
title: "Trajectory-construction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trajectory-construction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# 1 Data preprocessing

## 1.1 set R library loaction and load packages

```{r eval=TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
.libPaths('~/R/x86_64-pc-linux-gnu-library/4.3/')
library(ggplot2)
library(devtools)
library(tidyverse)
library(pheatmap)
```

## 1.2 load LineageNet

```{r eval=TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
library(LineageNet)
```

## 1.3 set color

```{r eval=TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
my_pal_science6 = c("#A9CA70","#C5D6F0","#F18C54",
                    "#9B59B6","#F7CAC9","#76D7C4")
```

## 1.4 load test data
We give the processed snRNA-seq data for E10 and E11 (see [Data-preprocessing-and-cell-annotation for processing](./vignettes/Data-preprocessing-and-cell-annotation.Rmd)) for Wot analysis
```{r eval=TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
data("bad_genes")
seob_E10_Mand_flow = readRDS(file = "/mnt/data/home/tycloud/workspace/scATAC/data/E10_Mand_WT_CNCC_Flow_tmp_scRNA_seq.rds")
seob_E11_Mand_flow = readRDS(file = "/mnt/data/home/tycloud/workspace/scATAC/data/E11_Mand_WT_CNCC_Flow_tmp_scRNA_seq.rds")


```


# 2 integrated
Wot analysis requires the downscaling of two temporal data that need to be inferred to the same space

```{r eval=TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
seob_E10_Mand_E11_Mand = RunSeurat(seob = list("E10_Mand"=seob_E10_Mand_flow,
                                               "E11_Mand"=seob_E11_Mand_flow),
                                   split_by="sample",
                                   integrated="RPCA",
                                   coln_add = "all",
                                   cellid_add = F,
                                   cellid=NULL,
                                   nfeatures=3000,
                                   outfeatures = bad_genes,
                                   npcs=20,
                                   min_batch_num = 100,
                                   resolution=0.2,
                                   k_anchor=1,
                                   k_score = 120,
                                   k_filter = 600,
                                   k_weight = 100,
                                   use_old_RNA = F,
                                   use_old_Seurat = F)
seob_E10_Mand_E11_Mand@meta.data$celltype_velocyto = str_c(seob_E10_Mand_E11_Mand@meta.data$sample,
                                                           seob_E10_Mand_E11_Mand@meta.data$celltype,
                                                           sep = "_")

seob_E10_Mand_E11_Mand@meta.data$time = ifelse(seob_E10_Mand_E11_Mand@meta.data$sample%in%"E10_Mand",1,2)
```
## 2.1 Check integration results
```{r eval=TRUE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
DimPlot(seob_E10_Mand_E11_Mand, reduction = "umap", group.by = "sample",label = T)
DimPlot(seob_E10_Mand_E11_Mand, reduction = "umap", group.by = "celltype_velocyto",label = T)
```

## 2.2 Save Seurat as an anndata object and use wot to complete trajectory inference
```{r eval=FALSE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
reticulate::use_condaenv("/mnt/data/home/tycloud/software/anaconda3/envs/scFates/")
Sys.setenv(RETICULATE_PYTHON = "/mnt/data/home/tycloud/software/anaconda3/envs/scFates/bin/python3")
seurat2anndata(obj = seob_E10_Mand_E11_Mand,
               outFile = "/mnt/data/home/tycloud/workspace/scATAC/data/E10_Mand_E11_Mand_WT_CNCC_SCT_Waddington_OT_scRNA_seq.h5ad",
               slot = 'data', main_layer = 'RNA',
               transfer_layers = NULL,
               drop_single_values = FALSE)
```
# 3 Completing trajectory construction with wot in python environment
Where Wot_path needs to replace the script we provided with your own path
```{r eval=FALSE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
python_path = "/mnt/data/home/tycloud/mambaforge/envs/cellrank3/bin/python"
Wot_path = "/mnt/data/home/tycloud/workspace/LineageNet/LineageNet/python/Wot.py"
input_file = "/mnt/data/home/tycloud/workspace/scATAC/data/E10_Mand_E11_Mand_WT_CNCC_SCT_Waddington_OT_scRNA_seq.h5ad"
out_file = "/mnt/data/home/tycloud/workspace/scATAC/data/Waddington_OT_E10_Mand_E11_Mand.csv"

execute_py(python_path = python_path,
           script_path = Wot_path,
           input_file=input_file,
           out_file=out_file)
```

## 3.1 plot reseult
```{r eval=FALSE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
Tab_Wot = read.csv(out_file)%>%column_to_rownames(var = "X")
pheatmap(Tab_Wot,
         cluster_rows = F,
         cluster_cols = F)
```

## 3.2 plot all time wot result
Here we provide test results for all adjacent period pairs of wot and visualize the
```{r eval=FALSE, echo=TRUE, message=TRUE, warning=TRUE, include=TRUE}
load(file = "/mnt/data/home/tycloud/workspace/scATAC/data/trajectory_DAG.RData")
plot_trajectory_DAG(Tab_l = Tab_list_Wot,
                    Tab_l2 = NULL,
                    node_df = cell_df,
                    color_by="color",
                    cellid_by="cellid",
                    scale_ = NULL,
                    cut_ = 0.0,
                    order_ = rev(cell_df$cellid),
                    node_size = 2,
                    x_size = 3,
                    y_size = 3,
                    width_top = 0.5,
                    alpha_min = 0.5,
                    width_min = 0.3,
                    width_max = 0.8,
                    edge_color = c("#136783"),
                    node_color = my_pal_science5)
```


